\documentclass[11pt]{article}

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{fullpage}
\usepackage{ifpdf}
\usepackage{deauthor, times}
\usepackage[usenames,dvipsnames]{color}
\usepackage{stmaryrd}
\usepackage[numbers]{natbib}
\usepackage{wrapfig}
\usepackage{textcomp}
\usepackage[breaklinks,draft=false,backref]{hyperref}

\ifpdf
  \usepackage{hyperref}
  \usepackage{graphicx}
\else
  \usepackage[dvips]{graphicx}
  \usepackage[dvips]{hyperref}
\fi

\newcounter{copyrightbox}

% Listings
%----------
\definecolor{darkgreen}{rgb}{0,0.6,0}
\lstloadlanguages{haskell}
\newcommand{\lsthaskell}{\lstset{
      language=haskell,
      %basicstyle=\ttfamily\ninett\footnotesize,
      flexiblecolumns=false,
			tabsize=2,
      %basewidth={0.5em,0.45em},
      %aboveskip={3pt},
      %belowskip={3pt},
      keywordstyle=\color{blue}\bfseries,
      commentstyle=\color{darkgreen}\itshape,
      morekeywords={foldl,fold},
			classoffset=1,
			upquote=true,
			keywordstyle=\color{Fuchsia}\bfseries,
			classoffset=0,
			mathescape=true,
      literate={+}{{$+$}}1 {/}{{$/$}}1 {*}{{$*$}}1 % {=}{{$=$}}1
               {>}{{$>$}}1 {<}{{$<$}}1
							 {dollar}{{\$}}1
               {\\\\}{{\char`\\\char`\\}}1
               {->}{{$\rightarrow$}}2 {>=}{{$\geq$}}2 {<-}{{$\leftarrow$}}2
               {<=}{{$\leq$}}2 {=>}{{$\Rightarrow$}}2
               {\ .}{{$\circ$}}2 {\ .\ }{{$\circ$}}2
               {>>}{{>>}}2 {>>=}{{>>=}}2 {=<<}{{=<<}}2
               {|}{{$\mid$}}1
							 {(-}{{$\in$}}1
						   {psi1}{{$\psi_1$}}1 {psi2}{{$\psi_2$}}1
							 {cup}{{$\cup$}}1
							 {cap}{{$\cap$}}1
							 {forall}{{$\forall$}}1
							 {vee}{{$\vee$}}1
							 {wedge}{{$\wedge$}}1
               {`member`}{{$\in$}}1
               {s.empty}{{\{\}}}1
               {leftbrace}{\{}1
               {rightbrace}{\}}1
               {profile0sing}{{ \{{\tt profile0}\}}}1
               {\$singleton\$startv}{{ \hspace{2.4em} \{{\tt startv}\}}}1
               {\$singleton\$n}{{  \{{\tt n}\}}}1
               {dotdotdot}{{$\ldots$}}3
    }}
\lstnewenvironment{codehaskell}
    { % \centering
			\lsthaskell
      \lstset{}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

% Formatting
%---------
\input{mydefs}
\newcommand{\C}[1]{\code{#1}}
\newcommand{\inang}[1]{\langle #1 \rangle}

% Formatting commands
% -------------------
\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\cf}[1]{{\small\tt #1}}
\newcommand{\rcf}[1]{\ensuremath{\mathrm{\cf{#1}}}}
\newcommand{\spc}[0]{\quad}
\newcommand{\rel}[1]{{R}_{\mathit{#1}}}
\newcommand{\conj}{~\wedge~}
\newcommand{\disj}{~\vee~}
\newcommand{\ilrulelabel}[1]{{\sc #1}}
\date{}
\begin{document}

\title{Representation without Taxation: A Uniform, Low-Overhead, and
High-Level Interface to Eventually Consistent Key-Value Stores}
%\subtitle{Research Summary}

\author{
  KC Sivaramakrishnan\thanks{This work was done at Purdue University, USA.} \\ 
  \small{University of Cambridge, UK}\\
  \small{{sk826@cl.cam.ac.uk}} 
  \and
  Gowtham Kaki \\ 
  \small{Purdue University, USA}\\
  \small{{gkaki@purdue.edu}} 
  \and
  Suresh Jagannathan\\
  \small{Purdue University, USA}\\
  \small{{suresh@cs.purdue.edu}} 
}


\maketitle

\begin{abstract}
Geo-distributed web applications often favor high availability over
strong consistency.  In response to this bias, modern-day replicated
data stores often eschew sequential consistency in favor of weaker
eventual consistency (EC) data semantics. While most operations
supported by a typical web application can be engineered, with
sufficient care, to function under EC, there are oftentimes critical
operations that require stronger consistency guarantees. A few
off-the-shelf eventually consistent key-value stores offer tunable
consistency levels to address the need for varying consistency
guarantees.  However, these consistency levels often have
poorly-defined \emph{ad hoc} semantics that is usually too low-level
from the perspective of an application to relate their guarantees to
invariants that must be respected by the application.  Moreover, these
guarantees are often defined in way that is strongly influenced by a
specific implementation of the data store.  While such low-level
implementation-dependent solutions do not readily cater to the
high-level requirements of an application, relying on ill-defined
guarantees additionally complicates the already hard task of reasoning
about application semantics under eventual consistency.

In this paper, we describe \name, a declarative programming model for
eventually consistent data stores. A novel aspect of \name is that it
abstracts the actual implementation of the data store via high-level
programming and system-level models that are agnostic to a specific
implementation of the data store.  By doing so, \name\ frees
application programmers from having to reason about their application
in terms of low-level implementation specific data store semantics.
Instead, programmers can now reason in terms of an abstract model of
the data store, and develop applications by defining and composing
high-level replicated data types. \name is equipped with a formal
specification language that is capable of expressing precise semantics
of high-level consistency guarantees (\emph{e.g.}, causal consistency)
in the abstract model.  Any eventually consistent key-value store can
support \name by implementing a thin shim layer and a choosen set of
high-level consistency guarantees on top of its existing low-level
interface.  We describe one such instantiation on top of Cassandra,
that includes support for causal and sequential consistency
guarantees, and coordination-free transactions. We present a case
study of a large web application benchmark to demonstrate \name's
practical utility.


\end{abstract}

\input{introduction}

\input{system-model}

\input{spec-lang}

\input{transactions}

\input{implementation}

\input{evaluation}

\input{conclusion}


{
%\small
\bibliographystyle{plainnat} \small \bibliography{all}
}

\end{document}

