\section{Introduction}
\label{sec:intro}

Eventual consistency facilitates high availability, but eventually consistent
replication leads to anomalies that have been
well-documented~\cite{Burckhardt2014, pldi15, Session, Dynamo, RedBlue}. While
applications are willing to tolerate most of these anomalies, there are some
that adversely effect the user experience, hence need to be avoided. For
instance, a social network application can tolerate out-of-order delivery of
unrelated posts, but causally related posts need to be delivered in the causal
order; a comment cannot be delivered before the post itself. The view count of
a video on Youtube need not necessarily reflect the actual count, but it
shouldn't appear to be decreasing. A bank account application may not always
show the accurate balance in an account, but neither should it let the balance
go below zero, nor should it display a negative balance. Bare eventual
consistency is often too weak to ensure such high-level application
invariants; stronger consistency guarantees are needed. To help applications
enforce such high-level invariants, off-the-shelf replicated data stores, such
as Cassandra and Riak, offer tunable consistency levels on per-operation
basis: applications can specify the consistency level for every read and write
operation they perform on the data store. However, consistency levels offered
by the off-the-shelf stores are often defined at a very low level. For
example, consistency levels in Cassandra and Riak assume the values of
\C{ONE}, \C{TWO}, \C{QUORUM}, \C{ALL} etc., desribing how many nodes in the
distributed system of the store must respond before a read or write operation
is declared success. It is not immediately apparent what permutation of these
low-level consistency guarantees would let the application enforce its
high-level level invariants. For instance, what should be the consistency
level of reads and writes to the \C{posts} table\footnote{We use the word
\emph{table} as an all-encompassing term for various key-value abstractions
provided by data stores.} so as to guarantee the causal order delivery of
posts in the aforementioned social network application? 

Furthermore, the semantics of low-level consistency guarantees are not
uniform across stores. For instance, while \C{QUORUM} means
\emph{strict quorum} (i.e., Lamport's quorum~\cite{LamportQuorum}) in
case of Cassandra, it means a \emph{sloppy quorum}~\cite{Dynamo} in
Riak. Complicating the matters further, the semantics are often
imprecisely, or even inaccurately, defined in the informal
vendor-hosted documentations of data stores. For instance, Datastax's
Cassandra documentation~\cite{dxlwt} claims that one can achieve
``strong consistency'' with ``quorum reads and writes'' in Cassandra.
While this claim appears reasonable superficially (because a pair of
quorum operations are serialized at one node, at least), it is
incomplete, at best, and inaccurate otherwise\footnote{The devil is in
the details of the timestamp-based last-writer-wins conflict
resolution strategy in Cassandra, which need not necessarily pick the
last writer due to the inevitable clock drift across
nodes.~\cite{TyconCassandra} and~\cite{JepsenCassandra} present
counterexamples and a more accurate claim.}. Another example of a
low-level consistency enforcement construct with vaguely defined
semantics is Cassandra's Compare-and-Set (CAS) operation, which is
advertised as a ``lightweight transaction'' and exposed as a
conditional write query (eg., \C{INSERT INTO users VALUES … IF NOT
EXISTS}).  Addition of CAS to Cassandra was coupled with the
introduction of a new consistency level named \C{SERIAL}.  Strangely,
\C{SERIAL} is not a valid query-level consistency parameter for a
write (conditional or not), while the rest (eg., \C{ONE}) are
valid\footnote{Given the advertised use cases for lightweight
transactions (such as maintaining uniqueness of usernames), one might
expect a CAS to be SERIAL by default. It is therefore unintuitive that
CAS accepts a consistency parameter, at least to the developers of
cassandra-cql, a popular Haskell library for programming with
Cassandra, whose API for CAS operation incorrectly hardcodes the
parameter to SERIAL. This bug has been reported and fixed.}.
Furthermore, Cassandra accepts a new \emph{protocol-level} consistency
parameter for a CAS operation (can be set to SERIAL), and its informal
description doesn't explain how this parameter interacts with the
query-level consistency parameter.  The only way to unravel this
complexity is to understand the nitty-gritty of the Paxos-based
implementation of CAS in Cassandra. In this quagmire of low-level
implementation details, it is easy to lose track of the original
intent: to obtain the high-level guarantees required by the
application.  

As applications become more complex, the default abstraction offered
by the most replicated data stores, namely, a key-value map with
simple conflict resolution strategy (eg., last-writer-wins), is no
longer enough. For instance, a highly available shopping cart
application, such as Amazon's, cannot rely on an LWW key-value store
to store the current contents of a shared shopping cart, because LWW
strategy always discards all but one of the concurrent updates to the
cart. Instead, what the application needs is a replicated \C{set} data
type, with a nuanced conflict resolution strategy that, for example,
merges non-conflicting updates using set union, and reconciles
conflicting updates by prioritizing additions over deletions. The need
for eventually consistent replicated data types with non-trivial
conflict resolution strategies has been felt in the context of many
other applications, that it spurred a whole new line of research on
Conflict-free Replicated Data Types (CRDTs)~\cite{crdts}. There are
now data stores that complement their default key-value abstraction
with CRDTs. The 2.0 release of Riak, for example, supports five
different CRDTs - a flag, register, counter, set, and a map.  However,
as acknowledged by Riak's own documentation on CRDTs, there is a vast
and ever-growing body of literature proposing new CRDTs to suit
various application requirements. While it is practically impossible
to keep up with the pace of developments in CRDT research, it is also
a bad software engineering practice to implement application-specific
semantics at the database layer. If the wide-spread adoption of
Ruby-on-Rails is any indicator, then programmers are also averse to
relying on vendor-specific features at the database layer, because it
adversely affects the portability of their
application~\cite{BailisSIGMOD15}. The other alternative  is for the
programmers to implement  the CRDTs they require in their application
on top of the key-value interface provided by the EC store. Indeed,
this is a hard engineering problem, as evident from the experience of
engineers at Instagram~\cite{BransonC2013}, who, at the expense of few
man-months, have implemented an op-based AWOR set CRDT on top of
Cassandra's wide rows. Moreover, this approach results in solutions
that are strongly coupled with semantics of the application and the
data store, hence are not reusable across applications and data
stores.

In this paper, we describe \name, a declarative programming framework
for eventually consistent data stores that was build to address the
issues discussed above. \name can be realized as a thin layer on top
of any off-the-shelf eventually consistent key-value store, and as
such provides a uniform implementation-independent interface to the
store. \name programmers reason in terms of an abstract model of the
ECDS, and any functionality offered by the store in addition to bare
eventual consistency, including stronger consistency guarantees,
transactions with tunable isolation levels etc., is required to have
well-defined semantics in the abstract model. We show that various
high-level consistency guarantees (eg., causal consistency) and
various well-known isolation levels for transactions (eg., read
committed) indeed have well-defined semantics in the abstract model.
\name is equipped with an expressive specification language that lets
data store developers succintly describe the semantics of the
functionality they offer. A similar specification language is exposed
to the application programmers, who declare the consistency
requirements of their application as specifications. Specifications
are constructed using primitive consistency relations such as
\emph{visibility} and \emph{session order} along with standard logical
and relational operators. A novel aspect of \name is that it can
compare the specifications written by application programmers and data
store developers, and automatically map application's requirements to
the appropriate store-level guarantees. Consequently, \name
programmers can write portable code that automatically adapts to any
data store that exposes its functionality via \name. 

Another key advantage of \name is that it lets application programmers
define new replicated data types without they having to be explicitly
supported by the underlying data store. For instance, all the five
CRDTs that are currently supported by Riak can be succintly
implemented in \name at the application layer, and the implementations
continue to work over Cassandra, which does not natively support
CRDTs. Moreover, operations over CRDTs in \name can be executed under
stronger consistency guarantees, which is currently not possible with
their native counterparts in Riak. We have implemented \name as a
shallow extension of Haskell, supported by a transparent shim layer on
top of Cassandra. Our implementation supports coordination-free
transactions over arbitrary data types, thus allowing programmers to
compose various CRDTs. We have implemented various CRDTs, including
the AWOR-set CRDT mentioned in the context of Instagram previously, in
\name. None of our implementations require more than 100LOC,
demonstrating the power of abstractions offered by \name. We have
conducted a variety of experiments with our implementations of CRDTs
to study their performance, and the results are summarized in
\S.~\ref{sec:experiments}.

Rest of this paper is organized as following. The next section
describes the abstract system model. The following section describes a
sample application to demonstrate the style of programming promoted by
\name. We also cover the anomalies exhibited by the application under
eventual consistency, which highlights the need for stronger
consistency guarantees with well-defined high-level semantics.
~\S~\ref{sec:language} introduces \name's specification language, and
presents the specifications of various high-level consistency
guarantees. 
